---
title: "TidyTemplate"
date: 2022-08-24
output: html_document
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
theme_set(theme_light())
library(lubridate)


```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2022-08-23")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```




# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

https://chip-dataset.vercel.app/

```{r Wrangle}

raw_chips <- read_csv("chip_dataset.csv") 

chips <- raw_chips|> 
  janitor::clean_names() |> 
  select (-x1) |> 
  mutate(release_date = as.Date(release_date),
         year = year(release_date),
         transistors = transistors_million * 1e6)
chips
```

```{r}
#how many cpus and gpus
chips |> count(type)
```

```{r}
chips |> 
  count(year = lubridate::year(release_date)) |> 
  ggplot(aes(year,n)) + 
  geom_col()
```
```{r}
chips |> 
  count(year = lubridate::year(release_date), type) |> 
  ggplot(aes(year,n, fill = type)) + 
  geom_col() +
  labs(x = "Year", 
       y = "# of chips")
```

```{r}
chips |> 
  group_by(year = lubridate::year(release_date)) |> 
  summarize(pct_gpu = mean(type == "GPU"), 
            median_transistors = median(transistors, na.rm = T), 
            geom_mean_transistors = exp(mean(log(transistors), na.rm = TRUE)))|> 
  ggplot(aes(year, pct_gpu))+
  geom_line()
  
```



```{r}
# lets look at the distribution of the transistors
chips |> 
  ggplot(aes(transistors)) +
  geom_histogram() +
  scale_x_log10(labels = comma_format())
```

```{r}
chips |> 
  ggplot(aes(transistors_million)) +
  geom_histogram() +
  scale_x_log10(labels = label_number(suffix = "M", big.mark = ","))
```


```{r}
# lets create a function 
summarize_chips <- function(tbl){
  tbl |> 
    summarize(pct_gpu = mean(type == "GPU"), 
            median_transistors = median(transistors, na.rm = T), 
            geom_mean_transistors = exp(mean(log(transistors), na.rm = TRUE)), 
            n = n(), 
  .groups = 'drop')|>  
            arrange(desc(n))
}
```


```{r}
chips |> 
  group_by(year = lubridate::year(release_date)) |> 
  summarize_chips() |> 
  ggplot(aes(year, median_transistors/1e6)) + 
  geom_line() +
  scale_y_log10()
```

```{r}
chips |> 
  ggplot(aes(release_date, transistors_million)) +
  geom_point() + 
  geom_smooth(method = "loess") +
  scale_y_log10()
```


```{r}
# doubling every two years/moore's laws i.e to double every two years, every year the log will go by .5
chips |> 
  mutate(year = year(release_date), moores_law = 2^(.5*(lubridate::year(release_date)-2000))) |> 
           distinct(year, moores_law) |> 
           arrange(year) 
  
```

```{r}
chips |>  
  mutate(year = year(release_date), 
         moores_law = 30*2^(.5*(year-2000))) |> 
  ggplot(aes(release_date, transistors_million)) +
  geom_point() + 
  geom_line(aes(y = moores_law), color = 'red')+
  geom_smooth(method = "loess") +
  scale_y_log10()
```
```{r}
chips |>  
  mutate(years_since_2000 = as.integer(release_date - as.Date("2000-01-01"))/365) |> 
  mutate(moores_law = 30*2^(.5*(years_since_2000))) |> 
  ggplot(aes(release_date, transistors_million)) +
  geom_point() + 
  geom_line(aes(y = moores_law), color = 'red')+
  geom_smooth(method = "loess") +
  scale_y_log10() +
  labs(x = "Chip release date", 
       y = "# of transistors (millioins)")
```

lets look at things by fountry i.e. something that makes chips

```{r}
chips |> 
  group_by(foundry) |> 
  summarize_chips()
```

```{r}
chips |> 
  group_by(foundry = fct_lump(foundry, 6)) |> 
  summarize_chips()
```

```{r}
# lets summarize both by foundry and type

chips |> 
  group_by(foundry = fct_lump(foundry, 6), type) |> 
  summarize_chips()
```

```{r}
chips |> 
  group_by(foundry = fct_lump(foundry, 6), type) |> 
  summarize_chips() |> 
  mutate(foundry = fct_reorder(foundry, n, sum)) |> 
  ggplot(aes(n, foundry, fill = type)) + 
  geom_col() +
  labs(y = "Chip creator")
```

```{r}
# lets compare tsmc, intel and gf what is their average number of transistors by year

chips |> 
  group_by(foundry = fct_lump(foundry, 2), 
           year)|>   
  summarize_chips() |> 
  ggplot(aes(year, median_transistors, color = foundry)) + 
  geom_line() +
  scale_y_log10()
```

```{r}

chips |> 
  mutate(foundry = fct_lump(foundry, 2)) |> 
 ggplot(aes(release_date, transistors_million, color = foundry)) +
  geom_point(alpha = 0.1) + 
  
  geom_smooth(method = "lm") +
  scale_y_log10() +
  labs(x = "Chip release date", 
       y = "# of transistors (millioins)")
```


```{r}
chips |> 
  ggplot(aes(fp16_gflops)) +
  geom_histogram() + 
  scale_x_log10(labels = number_format())
```

are gigaflops going the same way as number of chips? lets have a look 

```{r}
chips |> 
  filter(!is.na(fp32_gflops)) |> 
   ggplot(aes(release_date, fp32_gflops)) +
  geom_point() + 
    geom_smooth(method = "lm") +
  scale_y_log10() +
  labs(x = "Chip release date", 
       y = "Gigaflops(fp64)")
```

within a timeframe what is the relationship between the number of transistors and gigaflops

```{r}
chips |> 
  filter(year == 2021) |> 
  ggplot(aes(transistors_million, fp32_gflops)) + 
  geom_point()
```
there is not a lot of data here so lets do  this 

```{r}
chips |> 
  filter(!is.na(fp32_gflops)) |> 
  ggplot(aes(transistors_million, fp32_gflops)) + 
  geom_point() +
  facet_wrap(~year)
```
some years dont have much data at all so lets do this

```{r}
chips |>
  filter(!is.na(fp32_gflops)) |>
  group_by(year) |> 
  filter(n() >= 50) |> 
  ggplot(aes(transistors_million, fp32_gflops)) + 
  geom_point() +
  facet_wrap(~year) +
  scale_x_log10() + 
  scale_y_log10()
```
so generally the transistors and fp32_gflops seems correlated

lets take about correlcation matrix

one problem with correlation metrics is that time is correlated with everythign and that means everything is goign to be correlated with everything. 


```{r}
chips |> 
  ggplot(aes(release_date, process_size_nm)) + 
  geom_point()+  
scale_y_log10()
```
so as we can see the processors are getting smaller year by year. 

Lets do some linera modelling here

```{r}
lm(log(fp64_gflops) ~log(transistors) + 
     log(process_size_nm) + 
     year, data = chips) |> 
  summary()
```
```{r}
lm(log(fp64_gflops) ~log(transistors),  
      data = chips) |> 
  summary()
```
# interactive visualizations: lets do a shiny app

Go to file open the shiny web otion, select single file see we have done in the app.R file. 

```{r}
# selecting only numerical columns

                                                                                                                                                                                                                    numeric_choices <- chips |> 
select_if(is.numeric) |> #now we dont want the year column so lets do select(-year, -transistors) |> # if we just need to see the column name
  colnames()
```



